---
title: "Movielens"
date: '`r Sys.time()`'
output:
  html_document:
    #code_folding: hide
    number_sections: yes
    toc: yes
    toc_depth: 3
---

```{r setup, include=FALSE}
library(scales)
library(tidyverse)
library(knitr)

# set plot theme
theme_set(theme_bw())
```

# Load and preview data

Read data from the `ratings.csv` file
```{r load-data}
ratings <- read_delim('C:/Users/ds3/Downloads/ml-10m/ml-10M100K/ratings.dat', delim = '::',
                    col_names = c('user_id','movie_id','rating','timestamp'))
```

Loaded `r format(object.size(ratings), units="Mb")` of ratings data, containing `r format(nrow(ratings), big.mark = ",")` ratings. Here's a preview:
```{r preview-data}
head(ratings) %>% kable()
```

# Summary statistics

```{r dist-ratings}
# plot the distribution of rating values https://speakerdeck.com/jhofman/modeling-social-data-lecture-2-introduction-to-counting?slide=26
View(ratings)
ratings %>% group_by(rating) %>% ggplot( aes(x = rating)) + geom_histogram(binwidth = 0.25) + xlab('Rating') + ylab('Number of Ratings')
```

## Per-movie stats

```{r aggregate-by-movie}
# aggregate ratings by movie, computing mean rating and number of ratings
# hint: use the n() function for easy counting within a group
summary_movies <- ratings %>% group_by(movie_id) %>% summarise(mean = mean(rating), count = n())
View(summary_movies)
```

```{r dist-movie-popularity}
# plot distribution of movie popularity (= number of ratings the movie received)
# hint: try scale_x_log10() for a logarithmic x axis
summary_movies %>% ggplot( aes(x = count)) + geom_histogram(binwidth = 0.20) + scale_x_log10() + xlab('Number of Ratings') +ylab('Number of Movies')
```

```{r dist-mean-ratings-by-movie}
# plot distribution of mean ratings by movie https://speakerdeck.com/jhofman/modeling-social-data-lecture-2-introduction-to-counting?slide=28
# hint: try geom_histogram and geom_density
summary_movies %>% ggplot(aes(x = mean)) + geom_density(fill = 'black') + xlab('Mean Ratings by Movie') + ylab('Density')
```

```{r cdf-movie-pop}
# rank movies by popularity (number of ratings) and compute the cdf, or fraction of all views covered by the top-k movies https://speakerdeck.com/jhofman/modeling-social-data-lecture-2-introduction-to-counting?slide=30
# hint: use dplyr's rank and arrange functions, and the base R sum and cumsum functions
# store the result in a new data frame so you can use it in creating figure 2 from the paper below
mutate(summary_movies, total = sum(count), cum_sum = cumsum(count), cdf = (cum_sum/total)) %>% arrange((cdf)) %>% mutate( rank = row_number(cdf)) %>% ggplot(aes(x = rank, y = cdf)) + geom_line(fill = 'black')
# plot the CDF of movie popularity

```


# Per-user stats

```{r aggregate-by-user}
# aggregate ratings by user, computing mean and number of ratings
View(ratings)
str(ratings)
rating_dt <- ratings %>% group_by(user_id) %>% summarize(mean = mean(rating), count = n()) |> arrange(desc(count))
view(rating_dt)
```

```{r dist-user-activity}
# plot distribution of user activity (= number of ratings the user made)
# hint: try a log scale here
rating_dt %>% ggplot(aes(x = count)) + geom_histogram() + scale_x_log10() + xlab('User ID') + ylab('Number of Ratings')

```

# Anatomy of the long tail

```{r long-tail}
# generate the equivalent of figure 2a of this paper:
# note: don't worry about the "null model" lines
# just do the solid lines and dotted line (optional)
# https://5harad.com/papers/long_tail.pdf

# Specifically, for the subset of users who rated at least 10 movies,
# produce a plot that shows the fraction of users satisfied (vertical
# axis) as a function of inventory size (horizontal axis). We will
# define "satisfied" as follows: an individual user is satisfied p% of
# the time at inventory of size k if at least p% of the movies they
# rated are contained in the top k most popular movies. As in the
# paper, produce one curve for the 100% user satisfaction level and
# another for 90%---do not, however, bother implementing the null
# model (shown in the dashed lines).

head(ratings)

ratings_movie <- ratings %>% group_by(movie_id) %>% summarise(count = n()) %>% arrange(desc(count)) %>%  mutate(rank_by_movies = row_number())

rating_user <- inner_join(ratings,ratings_movie, by = c('movie_id'))
head(rating_user)

lets_See <- rating_user %>% select(- 'timestamp') %>% select(- 'count') %>% select(- 'rating') %>% group_by(user_id) %>% mutate(rank = row_number(rank_by_movies))%>% summarize(weirdest = max(rank_by_movies))

library(ggplot2)

# 100% Satisfaction
# Define your function (no change needed here)
my_function <- function(lets_See, x){
  lets_See <- lets_See %>% mutate(is_satisfied = ifelse(weirdest <= x, 1, 0))
  sum(lets_See$is_satisfied) / nrow(lets_See)
}

# Assuming you have a dataframe 'lets_See' with a column 'weirdest'
# Create x values
x = seq(0, length(unique(ratings$movie_id)), 1)

# Apply the function over each x
y = sapply(x, function(val) my_function(lets_See, val))

# Plot
df <- data.frame(x = x, y = y)
head(df)
ggplot(df, aes(x = x, y = y)) + geom_line() + xlab('Inventory Size') + ylab('Percent of User Satisfied')

# 90% Satisfaction
# Define your function (no change needed here)
ratings_movie <- ratings %>% group_by(movie_id) %>% summarise(count = n()) %>% arrange(desc(count)) %>%  mutate(rank_by_movies = row_number())

rating_user <- inner_join(ratings,ratings_movie, by = c('movie_id'))
head(rating_user)

lets_See1 <- rating_user %>% select(- 'timestamp') %>% select(- 'count') %>% select(- 'rating') %>% group_by(user_id) %>% mutate(rank = row_number(rank_by_movies))%>% summarize(weirdest = quantile(rank_by_movies, 0.9))

library(ggplot2)

# Define your function (no change needed here)
my_function1 <- function(lets_See1, x){
  lets_See1 <- lets_See1 %>% mutate(is_satisfied = ifelse(weirdest <= x, 1, 0))
  sum(lets_See1$is_satisfied) / nrow(lets_See1)
}

# Assuming you have a dataframe 'lets_See' with a column 'weirdest'
# Create x values
x1 = seq(0, length(unique(ratings$movie_id)), 1)

# Apply the function over each x
y1 = sapply(x1, function(val) my_function1(lets_See1, val))

x2 = c(x,x1)
y2 = c(y,y1)

# Plot
df2 <- data.frame(x = x2, y = y2)
head(df2)
nrow(df2)
ggplot(df2, aes(x = x, y = y)) +
  geom_point(color = "black", size = 1) + 
  xlab('Inventory Size') +
  ylab('Percent of User Satisfied')